Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-07-05 paper/paper.tex
revise 6/7 and code

\begin{itemize}
  \item The components of the configurations are lifted as arguments of the evaluation
    function.
  \item The refunctionalized AAM collects all possible configurations, but our final target is an 
    abstract definitional interpreter that cares about final evaluation results.
    So starting from this section our abstract semantic artifact returns a set of final 
    values instead of collected configurations. 
  \item Configurations still play an important role in caching, where a cache
    is a mapping from configurations to sets of final values. The caching algorithm
    ensures termination of the analysis, as well as performs a sound over-approximation
    of all possible concrete reachable paths and values.
  \item We simplify the continuations with a nondeterministic operator @nd@ such 
    that the function @aeval@ only uses one layer of continuation.
\end{itemize}

@@

\subsubsection{Caches}
% why cache? problem of set of states.
Recall that the latent assumption of the work-list algorithm in classical AAM is that 
if we have seen a state @s@, it means we also have seen all the successors of state @s@.
The caching algorithm we adopt here replays the same assumption but in a big-step 
manner: if we have seen the configuration @c@, then it means we also have seen the 
values that are evaluated from the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.
The case class @Cache@ and its operations are defined as follow:

@@

The @Cache@ contains two maps @in@ and @out@ which are both mappings from configurations 
to sets of @VS@. The @in@ cache contains mappings from the previous iteration of evaluation, 
and the @out@ cache contains mappings after the current iteration of evaluation. 
Once we have evaluated a term to some values, we update the @out@ cache. In the next 
iteration, we will use the @out@ cache from the previous iteration as the @in@ cache.

\subsubsection{Cache-Passing Style}
%motivation
We can now transform our interpreter into cache-passing style. We begin by defining the 
return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
its own accumulating operation @++@ which defines as a union of @Set[VS]@s and a join of caches.
The join operation of caches is defined as joining its two store components.

@@

If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, 
and update the @out@ cache with values from the @in@ cache. This retrieval from the @in@ 
cache may return an empty set of values if there is no such mapping for this configuration 
from the previous iteration.
In the initial iteration, we first conservatively assume that all computations can diverge, 
and if not, we update its values in the cache after the evaluation. In terms of partial orders and 
lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
The @out@ cache of this iteration will be used as the @in@ cache for the next iteration.

\subsubsection{Nondeterminism Abstraction}
We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
the closure set, and it actually repeats the code that creates the new environment and 
store for the body expression of a closure.
Based on observation, we present a nondeterministic operator that abstracts this pattern:

\begin{lstlisting}
  def nd[T](ts: Iterable[T], acc: Ans, f: (T, Cache, Ans => Ans) => Ans, m: Ans => Ans): Ans =
    if (ts.isEmpty) m(acc)
    else f(ts.head, acc.cache, (ans: Ans) => nd(ts.tail, acc ++ ans, f, m))
\end{lstlisting}

Given a collection of elements of type @T@, an initial accumulator of type 
@Ans@, a function @f@ that works on element's type @T@ and takes a continuation 
of type @Ans=>Ans@, and a function @m@ that is also type @Ans=>Ans@, @nd@ chooses
an element in the collection and applies @f@ to it, accumulates the value 
(@acc++ans@ where the @ans@ is the result comes from @f@) and eventually applies 
function @m@ on the accumulated value when there is no element in the collection.
This function is also written in cache-passing style, and the current accumulated 
cache @acc.cache@ is exposed to @f@ when applied.

Our function @nd@ is a variant implementation of @fold@ in continuation-passing style,
and @f@ also has to be passed as a function that takes a continuation.

\subsubsection{Put Them Together}
Given the skeleton of @aeval@ in cache-and-continuation-passing style, we may now
use the @nd@ operator to fill in the details. The first case of match is that @e@
is a @let@ expression:

\begin{lstlisting}
  case Let(x, App(f, ae), e) =>
    val closures = atomicEval(f, env, store)
    nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache, clsnd) =>
      val Clos(Lam(v, body), c_env) = cls;   val vbaddr = allocBind(v, new_time)
      val new_cenv = c_env + (v -> vbaddr); val new_store = store.update(vbaddr, atomicEval(ae, env, store))
      aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
        nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (vs, bdvsscache, bdnd) =>
          val Val(vals, time, vsstore) = vs;  val baddr = allocBind(x, time)
          val new_env = env + (x -> baddr);  val new_store = vssstore.update(baddr, vals)
          aeval(e, new_env, new_store, time, bdvsscache, { case Ans(evss, ecache) =>
            /* evaluate e with new env and store, then transfer to the next body's value */
            bdnd(Ans(evss, ecache.outUpdate(config, evss)))
          })
        }, clsnd) /* transfer to the next target closure of f */
      })
    }, continue)
\end{lstlisting}

We invoke @nd@ twice and perform a nesting two-step, depth-first evaluation over 
possible values of @App(f, ae)@. The values of branches from a fork point will be
accumulated inside of @nd@, and then returned to the upper fork point.

The outer application of @nd@ evaluates over the set of possible target closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bdvss@ is a set of @VS@ objects which represent values
from multiple computation paths when evaluating this single @body@ expression.

We next consider the inner application of @nd@, which evaluates over the set of body's 
value/stores @bdvss@. For each @VS@, the time-stamp and store in @VS@ will be instantiated to
the next application of @aeval@. The inner application of @aeval@ evaluates @e@ with the
substituted environment and store, then returns a set of values of one computation path 
to its continuation @bdnd@, where the continuation will accumulate the result and 
transfer the evaluation to the next body value.
The meta-continuation of the inner @nd@ is @clsnd@, meaning that we need to transfer 
to the next target closure of @f@. The meta-continuation of the outer @nd@ is just 
@continue@, as what we normally would do when writing CPS interpreters.

Note that the cache is used in a monotonic way: each function call to @aeval@ or @nd@ 
is passed with the latest cache from the most recent continuation of @aeval@ or @nd@.

@@

\begin{lstlisting}
  val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
  def analyze(e: Expr) = {
    def iter(cache: Cache): Ans = {
      val Ans(vss, new_cache) = aeval(e, mtEnv, mtStore, mtTime, cache, { case Ans(vss, cache) => 
          Ans(vss, cache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss))
      })
      if (new_cache.out == new_cache.in) Ans(vss, new_cache)
      else iter(Cache(new_cache.out, Store[Config, VS](Map())))
    } iter(mtCache) }
\end{lstlisting}

% TODO: explain more on how caches work
Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
to find the least fixed-point of cache starting from an empty cache.
If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
fact was discovered during this iteration, then we have reached a fixed-point of cache that 
has over-approximated the concrete evaluations, and thus the result can be returned.
Otherwise, we install the latest @out@ cache to the @in@ cache position when initializing 
the next iteration; in the meanwhile, an empty cache will be used as the @out@ cache. 
Note that the initial continuation of calling @aeval@ updates the values for the initial 
configuration in the final cache.

Now, what we have obtained is an abstract interpreter written in continuation-passing style.
Though the top-level function @aeval@ has only one continuation, but internally it uses
a @nd@ operator which is written with two layers of continuations. Furthermore, we added
a cache along with the interpreter and @nd@ operator.

@@

In the previous section, we presented a refunctionalized AAM utilizing the @nd@ operator
in extended continuation-passing style.
To obtain a definitional abstract interpreter in direct-style, we have multiple choices 
of how to proceed:
\begin{itemize}
  \item By replacing the continuations with monads \cite{filinski1994representing}, 
    we obtain the abstract definitional interpreters in monadic style which are similar to what 
    \citet{darais2017abstracting} describe.
  \item By representing the extended continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
    This transformation is the left-inverse of the CPS transform \cite{DBLP:journals/scp/Danvy94, DBLP:conf/lfp/DanvyL92}.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \Citet{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates.
    In this case, nondeterminism can be easily handled via @for@ comprehensions. 
\end{itemize}

@@

As previously identified, there are two layers of continuations: one for the normal stack 
structure of the analyzed language, and the other for the nondeterministic choices of closures.
This corresponds to the extended continuation-passing style.
But in our defining language (Scala), the delimited control operators rely on the continuation-passing 
transformation which is only one layer.

To address this problem, we reintroduce two operators @nd@ and @choices@ for handling 
nondeterministic choices in a way that is direct-style but mimics extended 
continuation-passing style. \\

\textit{Nondeterminism Operators, Again.}
%Introducing @Cache@ explicitly as the part of arguments of continuation @k@ reveals 
%that cache should be monotonically accumulated to the rest of computation. 
The first operator @nd@ is adapted from the previous one, but the meta-continuation
@m@ has been eliminated here. Now the only explicit continuation @k@ applies on tuple 
type @(T, Cache)@ where the value of type @T@ comes from the collection, and @nd@ performs 
recursive invocation on itself over elements in the collection @ts@.

\begin{lstlisting}
  def nd[T](ts: Iterable[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
    if (ts.isEmpty) acc
    else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
  }
\end{lstlisting}

The second operator @choices@ is implemented with the delimited control operator 
@shift@ but simply calls our first operator @nd@ in its body.

\begin{lstlisting}
  def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift {
    f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
  }
\end{lstlisting}

The function @choices@ takes two arguments: a set of elements of type @T@, and an initial cache.
The return type of @choices@ is @(T, Cache)@ with an annotation \verb|@cps[Ans]|, which 
denotes that @choices@ will iteratively return an element of type @T@ from the collection
along with the latest cache, but the final returned value of the continuation is type @Ans@.
The function @f@ is the delimited continuation captured by the @shift@ operator, which
is formed by the computation will be executed after the @choices@'s call-site.

Now we can use these two operators to redefine @aeval@. The abstract evaluator @aeval@ is
rewritten in direct-style, so that the last argument @cont@ disappears; we also add an 
annotation \verb|@cps[Ans]| on the return type to enable using delimited control operators 
inside of it.
Then, for every nondeterministic choice in the abstract interpretation, we can simply call 
@choices@ on the set of possible choices and write the program sequentially as concrete 
interpreters, and the subsequent statements and expressions become the delimited continuation
@f@. We have two invocations of @choices@, which correspond to the two invocations of @nd@ in
the refunctionalized form.
The function @choices@ also returns the latest cache to its left-hand side when invoked.
The cache will be still accumulated to the next call to @choices@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

\begin{lstlisting}
  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
    val config = Config(e, env, store, time)
    if (cache.outContains(config)) Ans(cache.outGet(config), cache)
    else {
      val new_time = config.tick; val new_cache = cache.outUpdate(config, cache.inGet(config))
      e match {
        case Let(x, App(f, ae), e) =>
          val closures = atomicEval(f, env, store)
          val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
          val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
          val new_store = store.update(vbaddr, atomicEval(ae, env, store))
          val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
          val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
          val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
          val new_vsstore = vsstore.update(baddr, vals)
          val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
          Ans(evss, ecache.outUpdate(config, evss))
        case ae if isAtomic(ae) =>
          val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
          Ans(vss, new_cache.outUpdate(config, vss))
      }
    }
  }
\end{lstlisting}

The function @analyze@ is also changed by calling the @reset@ operator around the function 
call of @aeval@ to set the delimiter. The afterward updating on cache becomes sequential.

\begin{lstlisting}
  def analyze(e: Expr) = {
    def iter(cache: Cache): Ans = {
      val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
      val new_cache = anscache.outUpdate(Config(e, mtEnv, mtStore, mtTime), vss)
      if (new_cache.out == new_cache.in) Ans(vss, new_cache)
      else iter(Cache(new_cache.out, Store[Config, VS](Map())))
    } iter(mtCache) }
\end{lstlisting}

Now we have arrived at the end of this series of transformations. 
The inverse transformation of the last step is basically CPS transforming the function @aeval@ and @nd@.
We may first explicitly expose the continuation at the level of @aeval@, and then convert @nd@ 
to extended continuation-passing style. Besides, these continuations must be manifestly embedded
at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.

>>>>

\begin{itemize}
  \item The components of the configuration are lifted as arguments of the evaluation
    function.
  \item The refunctionalized AAM collects all possible configurations, but our final target is an 
    abstract definitional interpreter that cares about final evaluation results.
    So starting from this section our abstract semantic artifact returns a set of final 
    values instead of collected configurations. 
  \item Configurations still play an important role in caching, where a cache
    is a mapping from configurations to sets of final values. The caching algorithm
    ensures termination of the analysis, as well as performs a sound over-approximation
    of all possible values and reachable paths.
  \item We simplify the continuations with a nondeterministic operator @nd@ such 
    that the function @aeval@ only uses one layer of continuation.
\end{itemize}

@@

\subsubsection{Cache-Passing Style}
% why cache? problem of set of states.
Recall that the latent assumption of the work-list algorithm in classical AAM is that 
if we have seen a state @s@, it means we also have seen all the successors of state @s@.
The caching algorithm we adopt here replays the same assumption but in a big-step 
manner: if we have seen the configuration @c@, then it means we also have seen the 
values that are evaluated from the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.
The case class @Cache@ and its operations are defined as follow:

@@

The @Cache@ contains two maps @in@ and @out@ which are both mappings from configurations 
to sets of @VS@. The @in@ cache contains mappings from the previous iteration of evaluation, 
and the @out@ cache contains mappings after the current iteration of evaluation. 
Once we have evaluated a term to some values, we update the @out@ cache. In the next 
iteration, we will use the @out@ cache from the previous iteration as the @in@ cache.
The iteration terminates when the @out@ cache is identical to the @in@ cache.

We can now transform our abstract interpreter into cache-passing style. We begin by defining the 
return type @Ans@ as a case class which contains a @VS@ set and a cache. @Ans@ also implements 
its own accumulating operation @++@ which defines as a union of @Set[VS]@s and a join of caches.
The join operation of caches is defined as joining its two store components.

@@

If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, 
and update the @out@ cache with the values from the @in@ cache. This retrieval from the @in@ 
cache may return an empty set of values if there is no such mapping for this configuration 
from the previous iteration.
In the initial iteration, we first conservatively assume that all computations can diverge, 
and if not, we update its values in the cache after the evaluation. In terms of partial orders and 
lattices, it is the case that starts the Kleene iteration from the bottom of the lattice.
The @out@ cache of this iteration will be used as the @in@ cache for the next iteration.

\subsubsection{Nondeterminism Abstraction}

We observe that in the refunctionalized AAM, @makeMCont@ is invoked with the tail part of
the closure set, and it actually repeats the code that creates the new environment and 
store for the body expression of a closure.
Based on observation, we present a nondeterministic operator that abstracts this pattern:

\begin{lstlisting}
  def nd[T](ts: Iterable[T], acc: Ans, k: (T, Cache) => Ans): Ans =
    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
\end{lstlisting}

Given a collection of elements of type @T@, an initial accumulator of type @Ans@,
a function @k@ that works on elements of type @T@, @nd@ chooses an element in
the collection and applies @k@ to it iteratively, and accumulates the value that
returned from @k@ when recuring on the rest of the collection. Eventually, @nd@
returns the accumulated value @acc@ when there is no element in the collection.
The latest accumulated cache @acc.cache@ is exposed to @k@ when applied.

\subsubsection{Put Them Together}

With the skeleton of cache-and-continuation-passing style @aeval@ in mind, we may now
use the @nd@ operator to fill in the details. The first case of match is that @e@
is a @let@ expression:

\begin{lstlisting}
  case Let(x, App(f, ae), e) =>
    val closures = atomicEval(f, env, store)
    nd[Storable](closures, Ans(Set[VS](), new_cache), { case (cls, clscache) =>
      val Clos(Lam(v, body), c_env) = cls;   val vbaddr = allocBind(v, new_time)
      val new_cenv = c_env + (v -> vbaddr); val new_store = store.update(vbaddr, atomicEval(ae, env, store))
      aeval(body, new_cenv, new_store, new_time, clscache, { case Ans(bdvss, bdcache) =>
        nd[VS](bdvss, Ans(Set[VS](), bdcache), { case (vs, bdvsscache) =>
          val Val(vals, time, vsstore) = vs;  val baddr = allocBind(x, time)
          val new_env = env + (x -> baddr);  val new_store = vssstore.update(baddr, vals)
          aeval(e, new_env, new_store, time, bdvsscache, { case Ans(evss, ecache) => 
            continue(Ans(evss, ecache.outUpdate(config, evss)))
          }) }) }) })
\end{lstlisting}

We invoke @nd@ twice and perform a nesting two-step, depth-first evaluation over 
possible values of @App(f, ae)@. 
The outer application of @nd@ evaluates over the set of possible target closures of
@f@. For each such closure, we go into its body with a new environment and a new store.
Remember that the function @aeval@ returns a set of @VS@ objects to its continuation,
so the continuation argument @bdvss@ represents a set of values from multiple computation 
paths when evaluating this single @body@ expression.

We next consider the inner application of @nd@, which evaluates over the set of body's 
value/stores @bdvss@. For each @VS@, the time-stamp and the store in @VS@ will be 
instantiated to the next application of @aeval@. The environment @new_env@ is built on
the outer environment variable @env@, which is the environment of this @let@ expression.
The inner application of @aeval@ evaluates @e@ with the substituted environment and store,
then returns a set of values of one computation path to its continuation, where we
apply the @continue@ function with these values and the updated cache. 
Note that the cache is used in a monotonic way: each function call to @aeval@ or @nd@ 
is passed with the latest cache from the most recent continuation of @aeval@ or @nd@.

@@

\begin{lstlisting}
  val mtCache = Cache(Store[Config, VS](Map()), Store[Config, VS](Map()))
  def analyze(e: Expr) = { 
    def iter(cache: Cache): Ans = { 
      val Ans(vss, anscache) = aeval(e, mtEnv, mtStore, mtTime, cache, ans => ans)
      if (anscache.out == anscache.in) Ans(vss, anscache)
      else iter(Cache(anscache.out, Store[Config, VS](Map())))
    } iter(Cache.mtCache) }
\end{lstlisting}

Finally, the @analyze@ function (as the entrance of the analysis) does a looping iteration
to find the least fixed-point of cache starting from an empty cache.
If the @out@ cache of this iteration is equivalent to its @in@ cache, which means no more 
fact was discovered during this iteration, then we have reached a fixed-point of cache that 
has over-approximated the concrete evaluations, and thus the result can be returned.
Otherwise, we install the latest @out@ cache to the @in@ cache position when initializing 
the next iteration; in the meanwhile, an empty cache will be used as the @out@ cache.

Now, as we can see, by using the @nd@ operator which abstracts over the meta-continuation,
the evaluation function can be simplified to one layer continuation-passing style.
The use two layers of continuations are unchained to two functions that each of them has one
continuation.
Now the continuation of @aeval@ plays the same role as the one in a concrete CPS interpreter.
Furthermore, we added a cache along with the abstract interpreter and @nd@ operator.

@@

In the previous section, we presented a refunctionalized AAM in extended continuation-passing style,
and then simplified it by utilizing the @nd@ operator.
To obtain a definitional abstract interpreter in direct-style, we have multiple choices of how to proceed:

\begin{itemize}
  \item By replacing the continuations with monads \cite{filinski1994representing}, 
    we obtain the abstract definitional interpreters in monadic style which are similar to what 
    \citet{darais2017abstracting} describe.
  \item By representing the continuation-passing style with delimited control operators
    such as @shift@ and @reset@, we derive a new form of abstract interpreters in direct-style.
    This transformation is the left-inverse of the CPS transform \cite{DBLP:journals/scp/Danvy94, DBLP:conf/lfp/DanvyL92}.
  \item In fact, we may also just use the same caching algorithm but with side effects such as
    assignments and mutations to update the cache, and then achieve the same definitional 
    interpreter with pushdown control-flows. \Citet{friedman-ai} in a tutorial on abstract 
    interpreters use this style to model caches and updates.
    In this case, nondeterminism can be easily handled via @for@ comprehensions. 
\end{itemize}

@@

We first transform the function @aeval@ to direct-style by removing the additional 
continuation argument @continue@ and sequentializing the expressions in @aeval@.
To enable using delimited control operators inside of @aeval@ we also add an 
annotation \verb|@cps[Ans]| on the return type, which tells the compiler to CPS
transform this function.
Back to direct-style on the case where @e@ appears as an atomic expression is straightforward: 
we simply remove the application of @continue@. Now the abstract evaluator @aeval@ looks as follows:

\begin{lstlisting}
  def aeval(e: Expr, env: Env, store: BStore, time: Time, cache: Cache): Ans @cps[Ans] = {
    val config = Config(e, env, store, time)
    if (cache.outContains(config)) Ans(cache.outGet(config), cache)
    else {
      val new_time = config.tick; val new_cache = cache.outUpdate(config, cache.inGet(config))
      e match {
        case Let(x, App(f, ae), e) => ...
        case ae if isAtomic(ae) =>
          val vss = Set(VS(atomicEval(ae, env, store), new_time, store))
          Ans(vss, new_cache.outUpdate(config, vss))
        } } }
\end{lstlisting}

The case where @e@ is a @let@ expression is not that obvious. As previously identified, 
function @nd@ works on a unknown number of nondeterministic choices, and it takes a 
function @k@ that works for every element in the collection. Here we present another
function @choices@ that uses delimited control operator @shift@ to capture that ``@k@''
when the program is direct-style and pass it to @nd@.

\begin{lstlisting}
  def nd[T](ts: Iterable[T], acc: Ans, k: ((T, Cache)) => Ans): Ans = {
    if (ts.isEmpty) acc else nd(ts.tail, acc ++ k(ts.head, acc.cache), k)
  }
  def choices[T](ts: Iterable[T], cache: Cache): (T, Cache) @cps[Ans] = shift {
    f: (((T, Cache)) => Ans) => nd(ts, Ans(Set[VS](), cache), f)
  }
\end{lstlisting}

The function @choices@ takes two arguments: a set of elements of type @T@, and an initial cache.
The return type of @choices@ is @(T, Cache)@ with an annotation \verb|@cps[Ans]|, which 
denotes that @choices@ will iteratively return an element of type @T@ from the collection
along with the latest cache once @f@ is invoked, but the final returned value of the 
function is type @Ans@.
The function @f@ is the delimited continuation captured by the @shift@ operator, which
is formed by the computation will be executed after the @choices@'s call-site.
Once the function @f@ (which is the function @k@ in @nd@) is invoked and returned, 
it returns to its call-site where @nd@ will be invoked on the rest elements of the collection.
Now we can use this function to sequentialize the @let@ case.

\begin{lstlisting}
  case Let(x, App(f, ae), e) =>
    val closures = atomicEval(f, env, store)
    val (Clos(Lam(v, body), c_env), clscache) = choices[Storable](closures, new_cache)
    val vbaddr = allocBind(v, new_time); val new_cenv = c_env + (v -> vbaddr)
    val new_store = store.update(vbaddr, atomicEval(ae, env, store))
    val Ans(bdvss, bdcache) = aeval(body, new_cenv, new_store, new_time, clscache)
    val (VS(vals, time, vsstore), vscache) = choices[VS](bdvss, bdcache)
    val baddr = allocBind(x, time); val new_env = env + (x -> baddr)
    val new_vsstore = vsstore.update(baddr, vals)
    val Ans(evss, ecache) = aeval(e, new_env, new_vsstore, time, vscache)
    Ans(evss, ecache.outUpdate(config, evss))
\end{lstlisting}

For every nondeterministic choice in the abstract interpretation, we can simply call 
@choices@ on the set of possible choices and write the program sequentially as concrete 
interpreters, and the subsequent statements and expressions after the call-site become 
the delimited continuation @f@. We have two invocations of @choices@, which correspond 
to the two invocations of @nd@ in the refunctionalized form in the last section.
The function @choices@ also returns the latest cache to its left-hand side when invoked.
The cache will be still accumulated to the next call to @choices@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

The function @analyze@ is also changed by calling the @reset@ operator around the function 
call of @aeval@ to set the delimiter.

\begin{lstlisting}
  def analyze(e: Expr) = { 
    def iter(cache: Cache): Ans = { 
      val Ans(vss, anscache) = reset { aeval(e, mtEnv, mtStore, mtTime, cache) }
      if (anscache.out == anscache.in) Ans(vss, anscache)
      else iter(Cache(anscache.out, Store[Config, VS](Map())))
    } iter(Cache.mtCache) }
\end{lstlisting}

Now we have arrived at the end of this series of transformations. 
The inverse transformation of the last step is basically CPS transforming the function @aeval@ and
explicitly using function @nd@ instead of function @choices@. Then, these continuations must be 
manifestly embedded at the call-site of @nd@, instead of implicitly provided by the @shift@ operator.
